Topics Causal Inference

### 2024-08-17 documents/paper.tex
Compute wgan coverage for estimated CATE; add to paper.

\section{Simulation Results}
Throughout the simulations I use $N=10,000$ observations and $40$ simulation repetitions, corresponding to~\cite{wager2018estimation}.
This low number is chosen to make the problem computationally feasible.
I use $100$ trees in all causal forest estimators and a sub-sampling share of $0.2$.
These parameters are much smaller than in the original~\cite{wager2018estimation} to reduce the run-time of the simulation.
I still find very similar results in the replication parts, as outlined below.
I follow~\cite{wager2018estimation} and report the expected MSE and expected coverage for CATE estimation by evaluating the predictions of the estimator on a new sample, also of size $10,000$.
To give an indication of noisiness coming from both a low number of simulations and approximating expected MSE and coverage, I also report the standard deviation of simulation runs.
For generic machine learning simulations I use a standard random forest as the baseline learner for $b_0(Z) = \E[Y(0)|Z]$ and a causal forest with similar hyperparameters as above for estimating $s_0(Z)$.
When analysing group average treatment effects (GATES), I define five groups based on the treatment effect quintiles predicted by the ML proxy.
To approximate the true GATES induced by the proxy I estimate this on a large sample of size $N=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
I use a main, auxiliary splitshare of $0.5$ throughout the analysis.

\subsection{Results: WA Replication and WGAN}

\begin{table}
    \caption{Simulation Results: WA Replication and WGAN}
    \begin{threeparttable}
        \center
        \begin{tabular}{lccccccc}
            & Dimension & MSE (True) & SD & MSE (Pop) & SD & Coverage & SD \\
            \toprule
            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel D:\ Placebo}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp6.tex} \\
            \bottomrule
        \end{tabular}
        \begin{tablenotes}
            \small
            \item \textbf{Notes}: This table reports simulation results for~\cite{wager2018estimation}.
            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
            \textit{Dimension} refers to the number of covariates in the DGP\@. \textit{Standard} DGP refers to sampling directly from the DGP, while \textit{WGAN} DGP refers to sampling from a WGAN trained based on a sample from the given DGP\@.
            MSE and coverage are calculated based on a test sample of $10,000$ observations. MSE \textit{TRUE} evaluates against the true CATE function of the respective DGP, while MSE \textit{Pop} evaluates against a CATE function estimated based on a (large) population draw and only applies to the WGAN DGP\@.
            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
            For details see the text.
        \end{tablenotes}
    \end{threeparttable}
\end{table}

\subsection{Results: Generic Machine Learning}

\begin{table}
    \caption{Simulation Results: ML Generic --- Cate Estimation}
    \begin{threeparttable}
    \center
    \begin{tabular}{lcc}
        Dimension & MSE & SD \\
        \toprule
        \multicolumn{3}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp3.tex} \\
        \midrule

        \multicolumn{3}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp4.tex} \\
        \midrule

        \multicolumn{3}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp5.tex} \\
        \midrule

        \multicolumn{3}{l}{\textit{Panel D:\ Placebo}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp6.tex} \\
        \bottomrule
    \end{tabular}
    \begin{tablenotes}
        \small
        \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml}.
        In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
        Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
        \textit{Dimension} refers to the number of covariates in the DGP\@.
        MSE and coverage are calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
        Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
        For details see the text.
    \end{tablenotes}
\end{threeparttable}

@@

% Generic ML: BLP and GATES Coverage
\begin{table}
    \caption{Simulation Results: ML Generic --- BLP and GATES Coverage}
    \begin{threeparttable}
        \center
        \begin{tabular}{lccccccc}
            Dimension & \multicolumn{7}{c}{Coverage} \\
            & \multicolumn{2}{c}{Best Linear Predictor} & \multicolumn{5}{c}{Group Average Treatment Effects} \\
            & $\beta_1$ & $\beta_2$ & $\gamma_1$ & $\gamma_2$ & $\gamma_3$ & $\gamma_4$ & $\gamma_5$ \\
            \toprule
            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
            \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp3.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
            \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp4.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
            \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp5.tex} \\
            \midrule

@@

        \begin{tablenotes}
            \small
            \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml} studying coverage for the BLP and GATES parameters (five groups).
            In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
            \textit{Dimension} refers to the number of covariates in the DGP\@.
            MSE and coverage are calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
            For details see the text.
        \end{tablenotes}
    \end{threeparttable}
\end{table}

@@

    \centering
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp3.png}
         \caption{Original WA DGP 3}\label{fig_gates:dgp3}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp4.png}
         \caption{High-dimensional --- 4 Covariates}\label{fig_gates:dgp4}
     \end{subfigure}

     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp5.png}
         \caption{High-dimensional --- 8 Covariates}\label{fig_gates:dgp5}
     \end{subfigure}
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp6.png}
         \caption{Placebo}\label{fig_gates:dgp6}
     \end{subfigure}

@@

    \centering
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp3.png}
         \caption{Original WA DGP 3}\label{fig_clans:dgp3}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp4.png}
         \caption{High-dimensional --- 4 Covariates}\label{fig_clans:dgp4}
     \end{subfigure}

     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp5.png}
         \caption{High-dimensional --- 8 Covariates}\label{fig_clans:dgp5}
     \end{subfigure}
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp6.png}
         \caption{Placebo}\label{fig_clans:dgp6}
     \end{subfigure}

>>>>

\section{Simulation Results}
Throughout the simulations I use $N=10,000$ observations and $40$ simulation repetitions, corresponding to~\cite{wager2018estimation}.
This low number is chosen to make the problem computationally feasible.
I use $100$ trees in all causal forest estimators and a sub-sampling share of $0.2$.
These parameters are much smaller than in the original~\cite{wager2018estimation} to reduce the run-time of the simulation.
I still find very similar results in the replication parts, as outlined below.
I follow~\cite{wager2018estimation} and report the expected MSE and expected coverage for CATE estimation by evaluating the predictions of the estimator on a new sample, also of size $10,000$.
Throughout, I construct confidence intervals with $95\%$ nominal coverage.
To give an indication of noisiness coming from both a low number of simulations and approximating expected MSE and coverage, I also report the standard deviation of simulation runs.
For generic machine learning simulations I use a standard random forest as the baseline learner for $b_0(Z) = \E[Y(0)|Z]$ and a causal forest with similar hyperparameters as above for estimating $s_0(Z)$.
When analysing group average treatment effects (GATES), I define five groups based on the treatment effect quintiles predicted by the ML proxy.
To approximate the true GATES induced by the proxy I estimate this on a large sample of size $N=1,000,000$; as noted in the paper, if the size of the auxiliary sample diverges the randomness due to splitting and training the proxy can be ignored.
I use a main, auxiliary splitshare of $0.5$ throughout the analysis.

\subsection{Results: WA Replication and WGAN}
I report results for the causal forest approach in Table~\ref*{table:wa_rep_wgan}.
First, the estimates for MSE --- using the true CATE function --- in column 2 are very close to the original results in~\cite{wager2018estimation}.
I find a MSE for 0.02 for $d=10$ with an SD smaller than 0.01 while they find MSE estimates ranging from 0.02 to 0.03 for dimensions ranging from 2 to 8 and a similar SD\@.
Notably, when I increase the dimensions to 15 and 30, respectively, I find comparable MSE estimates.
This shows, that causal forests are able to perform well in a sparse setting where the subset of covariates that are pure noise is potentially large.
The fourth row reports in Panel A reports simulation results based on an intermediate WGAN generator.
The MSE is more than three times as large as the directly simulated DGP when using the true CATE to compute the MSE (\textit{MSE True}), but delivers a similar MSE estimate when using an estimated CATE based on a large population draw (\textit{MSE Pop}).
Hence, using a WGAN generated DGP can in principle work well for simulation purposes when true CATE has a low-dimensional structure.
Turning to coverage in the last columns, I find that coverage is slightly above nominal level for all dimensions, which is different to the original results reported in~\cite{wager2018estimation}, who report coverage levels far below nominal level (e.g. 0.73 for $d=8$).
% [TODO Wgan coverage level after new simulation run using estimated CATE.]

Panels B and C report results for a CATE with higher dimension. As expected the performance of the Causal Forest significantly deteriorates.
MSE significantly increases, in particular for Panel C where the CATE depends on 8 covariates.
Even with only 4 covariates, coverage is sinificantly below the nominal level ranging from 0.83 to 0.76. For 8 covariates coverage is below 0.5 for all dimensions.
The WGAN generator approach result in similar MSE estimates when using the estimated CATE to compute MSE (0.22 vs. 0.19 using the standard approach).
However, for the more complicated CATE in Panel C, the WGAN approach suggests an MSE that is far below the actual MSE using the standard approach (44.99 vs. 133.81).
This is consistent with the previous argument, that the generator cannot preserve the high-dimensional structure. Hence, the CATE estimated based on a large population draw will be ``too simple''.
For the remaining simulations using the generic machine learning approach I stick to the standard approach.
Lastly, Panel D reports result for a Placebo CATE which is zero everywhere. The MSE is small and coverage is above nominal level.
Hence I find no indication for the claim made in connection with Figure 1 in~\cite{chernozhukov2023genml}, that random forest machine learning estimators can easily suggest a heterogeneous CATE for a placebo DGP\@.
\footnote{They, however, used a random forest directly (i.e.\ with a different target function and without modifications like honesty). Also, they probably reported out-of-the-box estimates which, since random forests are meant to address problems with more than one covariate, are prone to overfitting in this case. Using a random forest with cross-validation easily avoids overfitting.}
\begin{table}
    \caption{Simulation Results: WA Replication and WGAN}\label{table:wa_rep_wgan}
    \begin{threeparttable}
        \center
        \begin{tabular}{lccccccc}
            & Dimension & MSE (True) & SD & MSE (Pop) & SD & Coverage & SD \\
            \toprule
            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3 --- 2 Covariate CATE}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariate CATE}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariate CATE}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel D:\ Placebo}} \\
            \input{../bld/wa_replication/tables/wa_replication_dgp6.tex} \\
            \bottomrule
        \end{tabular}
        \begin{tablenotes}
            \small
            \item \textbf{Notes}: This table reports simulation results for~\cite{wager2018estimation}.
            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
            \textit{Dimension} refers to the number of covariates in the DGP\@. \textit{Standard} DGP refers to sampling directly from the DGP, while \textit{WGAN} DGP refers to sampling from a WGAN trained based on a sample from the given DGP\@.
            MSE and coverage are calculated based on a test sample of $10,000$ observations. MSE \textit{TRUE} evaluates against the true CATE function of the respective DGP, while MSE \textit{Pop} evaluates against a CATE function estimated based on a (large) population draw and only applies to the WGAN DGP\@.
            Coverage for $95\%$ confidence intervals reported.
            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
            For details see the text.
        \end{tablenotes}
    \end{threeparttable}
\end{table}

\subsection{Results: Generic Machine Learning}
I next turn to simulation results for the generic machine learning approach.
In the auxiliary sample I use a random forest learner for the baseline conditional average $b_0(z)$ and a causal forest for the CATE $s_0(z)$.
Results are median aggregated over 10 sample splits. For example, I use the median of the 10 parameter estimate for the heterogeneity loading $\hat{\beta_2^s}, s=1,\ldots,10$.
I also use the trained model $s(z)$ from the median slope parameter to evaluate the performance, i.e.\ construct BLP predictions.

\paragraph*{MSE}
While this approach does not directly target the CATE, it is still useful to evaluate the MSE\@.
After all, if the MSE is justified by some decision-theoretic approach, we might wonder whether trading off model misspecification for estimation precision results in smaller loss.
The results in Table~\ref*{table:generic_ml_cate} show that the MSE for estimating the true CATE is significantly worse compared to the causal forest approach.
While the BLP is of course not linear in the original covaiates, it seems like the gain estimation precision might still be offset by the approximation.
This is expected for DGPs with a low-dimensional CATE in which causal forests perform well (Panels A and B), but I would have expected a relatively better performance in Panel C.
One potential concern is that I used too few sample splits (the paper suggests 250 instead of 10), but this was computationally not feasible at this point.
The generic ML does in fact perform better in the placebo setting in Panel D, where the MSE is only a tenth compared to that of the causal forest.

\begin{table}
    \caption{Simulation Results: Generic ML --- Cate Estimation}\label{table:generic_ml_cate}
    \begin{threeparttable}
    \center
    \begin{tabular}{lcc}
        Dimension & MSE & SD \\
        \toprule
        \multicolumn{3}{l}{\textit{Panel A:\ Original WA DGP 3 --- 2 Covariate CATE}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp3.tex} \\
        \midrule

        \multicolumn{3}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariate CATE}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp4.tex} \\
        \midrule

        \multicolumn{3}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariate CATE}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp5.tex} \\
        \midrule

        \multicolumn{3}{l}{\textit{Panel D:\ Placebo}} \\
        \input{../bld/generic_ml/tables/generic_ml_dgp6.tex} \\
        \bottomrule
    \end{tabular}
    \begin{tablenotes}
        \small
        \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml}.
        In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
        Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
        \textit{Dimension} refers to the number of covariates in the DGP\@.
        MSE is calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
        Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
        For details see the text.
    \end{tablenotes}
\end{threeparttable}

@@

% Generic ML: BLP and GATES Coverage
\begin{table}
    \caption{Simulation Results: Generic ML --- BLP and GATES Coverage}
    \begin{threeparttable}
        \center
        \begin{tabular}{lccccccc}
            Dimension & \multicolumn{7}{c}{Coverage} \\
            & \multicolumn{2}{c}{Best Linear Predictor} & \multicolumn{5}{c}{Group Average Treatment Effects} \\
            & $\beta_1$ & $\beta_2$ & $\gamma_1$ & $\gamma_2$ & $\gamma_3$ & $\gamma_4$ & $\gamma_5$ \\
            \toprule
            \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3 --- 2 Covariate CATE}} \\
            \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp3.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariate CATE}} \\
            \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp4.tex} \\
            \midrule

            \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariate CATE}} \\
            \input{../bld/generic_ml/tables/generic_ml_blp_gates_coverage_dgp5.tex} \\
            \midrule

@@

        \begin{tablenotes}
            \small
            \item \textbf{Notes}: This table reports simulation results for~\cite{chernozhukov2023genml} studying coverage for the BLP and GATES parameters (five groups).
            In the auxiliary sample the BCA is estimated by a random forest regressor, while the CATE is estimated by a causal forest.
            Results for $40$ simulation runs with number of observations $N=10,000$, number of trees $200$ and resampling share $0.2$.
            \textit{Dimension} refers to the number of covariates in the DGP\@.
            Coverage is calculated based on a test sample of $10,000$ observations. SD refers to the standard deviation of this quantity over the simulation runs.
            Coverage for $95\%$ confidence intervals reported.
            Panel A reports results for the third DGP in~\cite{wager2018estimation}; Panel B and C increase the heterogeneity of the CATE function to 4 and 8 covariates, respectively; Panel D reports results for a CATE of zero (placebo).
            For details see the text.
        \end{tablenotes}
    \end{threeparttable}
\end{table}

@@

    \centering
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp3.png}
         \caption{Original WA DGP 3 --- 2 Covariate CATE}\label{fig_gates:dgp3}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp4.png}
         \caption{High-dimensional --- 4 Covariate CATE}\label{fig_gates:dgp4}
     \end{subfigure}

     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp5.png}
         \caption{High-dimensional --- 8 Covariate CATE}\label{fig_gates:dgp5}
     \end{subfigure}
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_gates_dgp6.png}
         \caption{Placebo}\label{fig_gates:dgp6}
     \end{subfigure}

@@

    \centering
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp3.png}
         \caption{Original WA DGP 3 --- 2 Covariate CATE}\label{fig_clans:dgp3}
     \end{subfigure}
     \hfill
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp4.png}
         \caption{High-dimensional --- 4 Covariate CATE}\label{fig_clans:dgp4}
     \end{subfigure}

     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp5.png}
         \caption{High-dimensional --- 8 Covariate CATE}\label{fig_clans:dgp5}
     \end{subfigure}
     \begin{subfigure}[b]{0.475\textwidth}
         \centering
         \includegraphics[width=\textwidth]{../bld/generic_ml/plots/generic_ml_clans_dgp6.png}
         \caption{Placebo}\label{fig_clans:dgp6}
     \end{subfigure}
